---
title: "Case Study"
author: "Pratheek Eravelli & Kanav Bhatnagar"
date: "3/26/2022"
output: pdf_document
---

# Project Setup

We will use the following packages for this homework assignment.  We will also read in data from the csv file.

```{r setup}
library(ggplot2)
prostate = read.table('prostate.txt', header = FALSE)
```

Now, we add labels to the columns.

```{r adding_labels}
colnames(prostate) = c('ID', 'psalevel', 'cancervolume', 'prostateweight', 
                       'age', 'hyperplasia', 'svi', 'capsular', 'gleason')
head(prostate)
```

# Analysis

## Building the initial model

First, we build a regression model with the required variables, and look at the summary statistics.

```{r lm}
prostate_lm = lm(psalevel ~ cancervolume + age + hyperplasia + svi + capsular + gleason,
                 data = prostate)
summary(prostate_lm)
```

## Manual model selection

We can see in the summary for the linear model that capsular penetration, i.e. `capsular` is the least significant variable since it has the highest *p*-value, which is over 0.05. We fit another linear model by removing this variable.

```{r rm_var1}
prostate_lm_reduced_1 = lm(psalevel ~ cancervolume + age + hyperplasia + svi + gleason,
                         data = prostate)
summary(prostate_lm_reduced_1)
```

Again, we notice that patient age, i.e. `age` is the least significant since it has the highest *p*-value, which is over 0.05. We fit another linear model by removing this variable.

```{r rm_var2}
prostate_lm_reduced_2 = lm(psalevel ~ cancervolume + hyperplasia + svi + gleason,
                           data = prostate)
summary(prostate_lm_reduced_2)
```

Now, we notice that the amount of benign prostatic hyperplasia, i.e. `hyperplasia` is the least significant since it has the highest *p*-value, which is over 0.05. We fit another linear model by removing this variable.

```{r rm_var3}
prostate_lm_reduced_3 = lm(psalevel ~ cancervolume + svi + gleason,
                           data = prostate)
summary(prostate_lm_reduced_3)
```

Finally, we notice that the Gleason score, i.e. `gleason` is the least significant since it has the highest *p*-value, which is over 0.05. We fit another linear model by removing this variable.

```{r rm_var4}
prostate_lm_reduced_4 = lm(psalevel ~ cancervolume + svi,
                           data = prostate)
summary(prostate_lm_reduced_4)
```

We see that all the variables in the model have a *p*-value less than 0.05, which means that their slopes are all statistically significant, and they contribute to explaining the variation in the response.

Thus, the final model we have is `psalevel =` `r prostate_lm_reduced_4$coefficients[1]` + `r prostate_lm_reduced_4$coefficients[2]` $\cdot$ `cancervolume` + `r prostate_lm_reduced_4$coefficients[3]` $\cdot$ `svi`.


## Unusual observations


## Model assumptions

